Intro to Linear Regression

### 2018-11-18 DATA606/Intro to Linear Regression.Rmd
Work on Chapter 7 Assignment

(a) Describe the relationship between number of calories and amount of carbohydrates (in grams) that Starbucks food menu items contain.

(b) In this scenario, what are the explanatory and response variables?

(c) Why might we want to fit a regression line to these data?

(d) Do these data meet the conditions required for fitting a least squares line?

@@

(b) Interpret the slope and the intercept in this context.

(c) Calculate $R^2$ of the regression line for predicting height from shoulder girth, and interpret it in the context of the application.

(d) A randomly selected student from your class has a shoulder girth of 100 cm.  Predict the height of this student using the model.

(e) The student from part (d) is 160 cm tall. Calculate the residual, and explain what this residual means.

(f) A one year old has a shoulder girth of 56 cm. Would it be appropriate to use this linear model to predict the height of this child?

@@

(a) Write out the linear model.

(b) Interpret the intercept.

(c) Interpret the slope.

(d) Interpret $R^2$.

(e) Calculate the correlation coefficient.

>>>>

(a) Describe the relationship between number of calories and amount of carbohydrates (in grams) that Starbucks food menu items contain. **Positive and fairly linear.**

(b) In this scenario, what are the explanatory and response variables? **The response variable is calories.  The explanatory variable is the grams of carbohydrates.**

(c) Why might we want to fit a regression line to these data? **The data seem like they could be summarized well with a regression line.**

(d) Do these data meet the conditions required for fitting a least squares line? **Yes.  The residuals are normally distributed.  There is no pattern aparent in the residuals plot.  And the relationship looks linear as previously stated.**

@@

$\widehat{\text{height}}=\beta_0 + \beta_1\times\text{girth}$

$\beta_1 = \frac{s_\text{height}}{s_\text{girth}}R = \frac{9.41}{10.37}0.67 = 0.6079749$

$y - y_0 = \beta_1(x-x_0) \rightarrow y - 171.14 = 0.6079749 (x-107.20)$

$\beta_0 = 171.14 - (-107.2 \times 0.6079749) \rightarrow \beta_0 = 171.14 - 65.17491 \beta_0 = 105.9651$

$\widehat{\text{height}}=\beta_0 + \beta_1\times\text{girth} = 105.9651 + 0.6079749 \times \text{girth}$

(b) Interpret the slope and the intercept in this context. **The intercept (average height) is roughly 106 centimeters.  For every centimeter in sholder girth there is about 0.61 centimeters of height.**

(c) Calculate $R^2$ of the regression line for predicting height from shoulder girth, and interpret it in the context of the application.  $R^2 = 0.67 \times0.67 = 0.4489$
**This means that about 45% of the variability in height is explained by the sholder girth.**

(d) A randomly selected student from your class has a shoulder girth of 100 cm.  Predict the height of this student using the model. $\widehat{\text{height}}= 105.9651 + 0.6079749 \times 100 = 105.9651 + 60.79749 = 166.7626$

(e) The student from part (d) is 160 cm tall. Calculate the residual, and explain what this residual means. **The residual is -6.7626 cm which is the amount overestimated by the model.**

$\text{Residual} = \text{Observed} - \text{Expected} = 160 - 166.7626 = 6.7626$

(f) A one year old has a shoulder girth of 56 cm. Would it be appropriate to use this linear model to predict the height of this child? **I don't think so.  The model worked well for a student in part d that is 160 cm tall which is over 5 feet.  The model works for adults.  The relationship between these two variables is not likely to hold at such a young age.**

@@

(a) Write out the linear model. $\widehat{\text{heart}} = -0.357 + 4.034 \times \text{body}$

(b) Interpret the intercept. **The weight of the cat's heart if it weighed nothing.**

(c) Interpret the slope. **For every one kg of body weight there will be 4.034 grams of heart weight.**

(d) Interpret $R^2$. **Roughly 64% of the variation is explained by the cat's body weight.**

(e) Calculate the correlation coefficient. $R = \sqrt{R^2} = \sqrt{0.6466} = 0.8041144$
